Large scale risks from pesticides in small streams

### 2016-09-30 article.tex
Update on Overleaf.

\subsection{Pesticides in small water bodies}
% man kann sich natürlich fragen, warum hier der datensatz reduziert wird, wenn catchment size eh kaum einfluss hat. dafür braucht es oben eine gute begründung. auch warum alle analysen mit dem vollen datensatz gemacht werden, außer dieser
The dataset comprised 12,710 samples from 1,295 small water bodies.
In 1,173 samples (0.3\% of all and 5.6\% of samples with detects) RQ higher than 1 were observed.
Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ exceedances (Figure~\ref{fig:fig6}).
For Thiacloprid, Imidacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ~\textgreater~1. 
The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (29.5, 14.2 and 21.4 \% of all samples with detects).
The highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 82\% of all samples were it was analysed, see also Supplemental Figure~S9).
Glyphosate was the compound with the highest detection rates, followed by Boscalid and Isoproturon. 
However, only the latter showed RQ exceedances (Figure~\ref{fig:fig6}).
In 44.8\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S8). 
%exceedances per site sollten hier erwähnt werden, nicht erst in Diskussion

@@

%% -------------------------------------------------------------------------
\section{Discussion}
\subsection{Overview on the compiled dataset}
The compiled dataset of governmental monitoring data represents currently the most comprehensive one available for Germany.
Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
The data compiled and analysed here for Germany is of similar quantity and quality.

% current problems in monitoring and possible solutions
Nevertheless, a nationwide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
There are not only big differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), but also the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and differences in the quality of chemical analyses. 
For Thiacloprid, Imidaclorpid and Chlorpyrifos LOQ were above RAC.
For these compounds a lowering of LOQ is essential for reliable assessment.
Moreover, would a nation-wide assessment benefit from a harmonized spectrum of analysed compounds between federal states.

Given their high abundance in the landscape \citep{nadeau_hydrological_2007} SWB are underrepresented in the current monitoring (Figure~\ref{fig:fig3}). 
Especially, streams below 10~km\textsuperscript{2} are missing, which could be attributed to the missing categorisation in the WFD.

\subsection{Influence of agricultural land use and catchment size}
% agriculture
We found a strong influence of agriculture on the pollution of streams.
If there is more the 25\% agriculture within a catchment pesticides, it is likely that RAC will be exceeded with an increase in fully agricultural catchments (above 75 \% agriculture).
To our knowledge this is the first study investigating such thresholds of pesticide exposure.
Thresholds for agricultural land use have only been investigated in the past for biological communities.
\citet{feld_response_2013} found change points of biological community metric for agricultural land use at 40\% in low lands. 
Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\%.
Our results coincide with these thresholds and suggest that pesticides might contribute to the observed biological changes.

% size
We did not find a relationship between pesticide pollution and catchment size.
However, previous studies have shown that SWB are more polluted that bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
This could be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams being \textless $100~km^2$ (Figure~\ref{fig:fig3}, top).
For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
Another reason might be the unequal distribution of catchment sizes, with fewer sites below $10~km^2$ and above $100~km^2$ catchment size (Figure~\ref{fig:fig3}, top).

\subsection{Effect of precipitation on pesticide exposure}
Our results revealed that pesticide sampling for chemical monitoring in Germany is mainly performed when no precipitation occurs. 
Nevertheless, we found higher RQ if samples where taken after rainfall events. 
Samples taken at the day of a rainfall-event showed less risk, than samples taken one day after a rainfall-event.
This could be explained by a sampling just before the actual rainfall event.
Pesticide concentrations in agricultural SWB generally show short term peak concentrations \citep{wittmer_significance_2010}, therefore, a sampling at the day after a rainfall-event might miss peak concentrations.
The effects of precipitation were more pronounced for the probability to exceed LOQ, with smaller effect sizes for RQ.
This could be attributed to a higher variability of absolute concentrations.
Overall, our results indicate that current pesticide monitoring strongly underestimates pesticide exposure.
Automatic event-drive samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for SWB \citep{lorenz_specifics_2016}.

We found highest the probability of exceeding LOQ during summer (9.9\% for Q2) and lowest in the first Quarter (2.8\%, Figure~\ref{fig:fig5}, bottom right).
This yearly pattern coincides with their main application season for pesticides.
Nevertheless, there are compound specific differences in the yearly pattern, which explain the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
For example, the herbicide Diflufenican showed highest RQ during the winter quarters (Supplemental Table~S4), which is the application period it is registered for in Germany \citep{bvl_online_2016}.
Our study suggest that pesticide exposure shows compound specific spatio-temporal dynamics.
Currently, little is known on these and further research on those might provide useful information for future ecological risk assessment.

\subsection{Pesticides in small water bodies}
Our results suggest that SWB are frequently exposed to biologically relevant pesticide concentrations.
\citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances for organophosphate insecticides. 
By contrast, our results revealed that neonicotinoid insecticides show high exceedances, followed by the organophosphate chlorpyrifos. 
This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (between 1,290 and 2,113 samples in SWB, Figure~\ref{fig:fig6}). 
Our results shows that this particular class of insecticides may currently pose a high risk to freshwater ecosystems.
Compared to \citet{stehle_pesticide_2015} we found much lower rates RAC exceedances (0.3\% of 372,304 samples with RAC vs 44\% of 1,566 samples). 
This can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality, also during periods with lower pesticide usage and at natural sites. 
Contrary, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
This might reflect differences in pesticide use between countries and defined RACs.

From the definition of RAC it follows that if the concentration of a compound exceeds its RAC ecological effects are expected.
Accordingly, \citet{stehle_agricultural_2015} found that biological diversity is significantly reduced at a RQ of 1.12.
We found RQ values greater than 1.12 at 25\% of all SWB sites. 
Consequently, we conclude that pesticides from agricultural land use are a major treat to small water bodies, the biodiversity they host and the services they provide. 
Additional treat arises when taking into account that most pesticides do not occur individually but in mixtures \cite{schreiner_pesticide_2016} and may co-occur with other stressors \citep{schafer_contribution_2016}.

% Approval / Risk Assessment
Monitoring data, despite the outlined limitations, provides an opportunity to study large scale environmental occurrences of pesticides.
Nevertheless, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}. 
The high exceedances of RAC indicate that the approval process for pesticides must be checked and refined.

>>>>

\subsection{Pesticides in small water bodies}
% man kann sich natürlich fragen, warum hier der datensatz reduziert wird, wenn catchment size eh kaum einfluss hat. dafür braucht es oben eine gute begründung. auch warum alle analysen mit dem vollen datensatz gemacht werden, außer dieser
The dataset comprised 12,710 samples from 1,295 small water bodies.
In 1,173 samples (0.3\% of all and 5.6\% of samples with detects) RQ higher than 1 were observed.
Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ exceedances (Figure~\ref{fig:fig6}).
For Thiacloprid, Imidacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ~\textgreater~1. 
The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (29.5, 14.2 and 21.4 \% of all samples with detects).
The highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 82\% of all samples were it was analysed, see also Supplemental Figure~S9).
Glyphosate was the compound with the highest detection rates, followed by Boscalid and Isoproturon. 
However, only the latter showed RQ exceedances (Figure~\ref{fig:fig6}).
In 44.8\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S8). 
%exceedances per site sollten hier erwähnt werden, nicht erst in Diskussion
% und man könnte sich fragen, ob es da unterschiede zwischen den bundesländern gab

@@

%% -------------------------------------------------------------------------
\section{Discussion}
\subsection{Overview on the compiled dataset}
The compiled dataset of governmental monitoring data, with a particular focus on small streams, represents currently the most comprehensive one available for Germany.
Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
% wäre es nicht besser Stone 2014 ES&T für die USA zu nehmen? Der Link verweist ja gerade auf Daten die nicht aufbereitet sind usw.
The data compiled and analysed here for Germany is of similar quantity and quality.
%zumindest mit bezug auf die US Daten stimmt diese aussage nicht

% current problems in monitoring and possible solutions
Nevertheless, a nationwide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states.
% was wiederum die oben aufgeworfene frage verstärkt, ob es bei den ergebnissen unterschiede zwischen den staaten gab...
For Thiacloprid, Imidaclorpid and Chlorpyrifos the LOQs were above the RAC, which means that some exceedances were likely not detected.
For these compounds a lowering of LOQ is essential for reliable assessment.
Moreover, a nation-wide assessment benefit from a harmonized spectrum of analysed compounds between federal states.
% ok, aber dann musst du auch irgendwo positiver formulieren, worauf sich die ergebnisse von uns jetzt beziehen - schon mal nicht auf Deutschland...

Given their high abundance in the landscape \citep{nadeau_hydrological_2007} SWB are underrepresented in the current monitoring (Figure~\ref{fig:fig3}). 
Especially, sampling sites in streams below 10~km\textsuperscript{2} are disproportionally low, which may be attributed to the missing categorisation in the WFD.

\subsection{Influence of agricultural land use and catchment size}
% agriculture
We found a strong influence of agriculture on the pollution of streams.
If there is more the 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase in the likelihood of RAC exceedance in fully agricultural catchments (above 75 \% agriculture).
% likely can be defined (0.5 it seems?) 
To our knowledge this is the first study investigating such thresholds of pesticide risk.
Previous studies examined thresholds for \%agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
\citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use  in low land streams. 
Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\% agricultural land use for ....
%kurze beschreibung welche Gewässer und Region
Our results coincide with these thresholds and suggest that pesticides might contribute to the observed biological changes.

% size
We did not find a relationship between pesticide pollution and catchment size.
However, previous studies showed that SWB are more polluted that bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
This can be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams with catchments below \textless $100~km^2$ (Figure~\ref{fig:fig3}, top).
For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
Another reason might be the unequal distribution of catchment sizes, with fewer sites below $10~km^2$ and above $100~km^2$ catchment size (Figure~\ref{fig:fig3}, top).
%Argument nicht ganz klar - sample size beeinflusst pollution state? Ansonsten ist das doch das gleiche argument, dass gradient insgesamt nur 10-100 war...

\subsection{Effect of precipitation on pesticide exposure}
Our results revealed that most samples for pesticide monitoring in Germany are taken in the absence of prior precipitation events. Given that we found higher risk from pesticides (RQ) for samples taken after rainfall events.
%would be good to be more quantitative here - was meint higher
Interestingly, samples taken at the day of a rainfall-event showed a slightly lower risk, than samples taken one day after a rainfall-event. This can may be explained on the one hand by the sampling preceding the actual rainfall event on the sampling day, which could not be accounted for in analysis due to a lack of data on the hour of sampling and rain event.
On the other hand, the delay between the start of the precipitation and the peak concentrations in streams varies with context variables such as geology, riparian land cover and slope. Thus, even a sampling after the initiation of a precipitation event may not capture the short term peak concentrations \citep{wittmer_significance_2010}.
The effects of precipitation were more pronounced for the probability to exceed LOQ, with smaller effect sizes for RQ.
This could be attributed to a higher variability of absolute concentrations.
Overall, our results indicate that current pesticide monitoring relying on grab sampling, largely disconnected from precipitation events, underestimates pesticide risks.
Automatic event-drive samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for SWB \citep{lorenz_specifics_2016}. 
% ist auch die frage, wie häufig du wirklich SWB, was auch stehende gewässer inkludiert, meinst - unsere ergebnisse sind ja nur für streams - macht eigentlich mehr sinn dann auch small streams zu schreiben oder sowas, da gerade für standgewässer die sache wesentlich komplizierter ist

We found highest the probability of exceeding LOQ during summer (9.9\% for Q2) and lowest in the first Quarter (2.8\%, Figure~\ref{fig:fig5}, bottom right).
This yearly pattern coincides with their main application season for pesticides in Central Europe.
Nevertheless, the annual pattern differed between compounds, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
%etwas seltsam in den sätzen davor bezug auf LOQ und erklärung für RQ
For example, the herbicide Diflufenican showed highest RQ during the winter quarters (Supplemental Table~S4), which is the application period it is registered for in Germany \citep{bvl_online_2016}.
Our study suggests that pesticide risks display compound specific spatio-temporal dynamics.
Currently, little is known on these, but would provide useful information for pesticide monitoring and risk assessment, given that the sensitivity of organisms is often life stage dependent
%ref needed
and temporal dynamics would inform on concurrent exposure to multiple pesticides as well as to parameterise toxikokinetik and toxikodynamic models.
% (1) Ashauer, R.; Albert, C.; Augustine, S.; Cedergreen, N.; Charles, S.; Ducrot, V.; Focks, A.; Gabsi, F.; Gergs, A.; Goussen, B.; et al. Modelling survival: exposure pattern, species sensitivity and uncertainty. Scientific Reports 2016, 6, 29178.

\subsection{Pesticides in small water bodies}
Our results suggest that SWB are frequently exposed to ecologically relevant pesticide concentrations.
%siehe bemerkung oben bzgl SWB
\citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances for organophosphate insecticides. 
By contrast, our results that neonicotinoid insecticides showed highest exceedances of RACs, followed by the organophosphate chlorpyrifos. 
This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (between 1,290 and 2,113 samples in SWB, Figure~\ref{fig:fig6}). 
%unklar warum between - und SWB
Overall, our results suggest that neonicotonoids may currently pose a high risk to freshwater ecosystems.
% hier könnte man zumindest auch kurz andere Studien erwähnen, gab ja sehr viel zu neoniks in den letzten Jahren. Z.b. OUr results add to the growing literature on risks from neonicotinoids in ecosystems such as ...
Compared to \citet{stehle_pesticide_2015} we found much lower rates of RAC exceedances (0.3\% of 372,304 samples with RAC vs 44\% of 1,566 samples). 
This may be attributed to two differences: 1) Different objectives of governmental monitoring and scientific studies. Scientific studies may be slightly biased towards streams with high pollution to detect effects, whereas monitoring aims mainly at spatially representative surveillance of water quality, also during periods with lower pesticide usage and at natural sites. 2) DIfferent sampling 
% ich meine dass bei Stehle und Schulz das meiste event driven war - hier diesen zweiten punkt nennen

By contrast, \citet{knauer_pesticides_2016} analysed Suisse pesticide monitoring data and found exceedances mainly for herbicides and fungicides and for only one insecticide, Chlorpyrifos.
% hier auch unterschiede in % der samples?
This might reflect differences in pesticide use between countries and, albeit to a smaller extent, in the RAC values.
%dieses argument torpediert die generelle nutzung von RACs als indikator für Risiko! Wenn er variiert würde es bedeutetn, dass es keinen festen Bezug zum ökoloigischen Risiko gibt...

From the definition of RAC it follows that if the concentration of a compound exceeds its RAC ecological effects are expected.
% wie gesagt widerspricht der aussage davor - vielleicht das als grund für unterschied einfach weglassen?
Indeed, \citet{stehle_agricultural_2015} found that biological diversity of stream invertebrates was significantly reduced at a RQ of 1.12.
%give number - how much reduced
We found RQ values greater than 1.12 at 25\% of all SWB sites. 
%kannst du nicht einfach oben dann RQ 1 in den resultaten angeben? Zudem sie auch schon abnahme bei 1/10 RAC fanden! das könnte man hier besser darstellen und vielleicht hier zusätzlich geben, a RAC exceedance of 1/10 was observed at xx5 of sites.
Consequently, we conclude that agricultural pesticides are a major threat to small streams, the biodiversity they host and the services they provide. 
This threat may exacerbate because esticides often occur in mixtures \cite{schreiner_pesticide_2016} and co-occur with other stressors \citep{schafer_contribution_2016}.

% Approval / Risk Assessment
Monitoring data, despite the outlined limitations, provides an opportunity to study large scale environmental occurrence patterns of pesticides.
Nevertheless, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}, though the sampling design need to account for precipitation events to provide robust data. 
The high exceedances of RAC indicate that the authorisation process for pesticides may require further development and refinement.
